---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# easyicer

<!-- badges: start -->

<!-- badges: end -->

This R package contains the `summed_output()` function which takes a column as a character string storing a vector of numeric values within a dataframe and calculates the sum of the vector. The `icercalc()` and `icerplot()` functions can be used to calculate and plot ICER values based on inputs obtained from `summed_output()`. It is designed to help users automate repetitive sum operations based on user-defined variable groupings and filter conditions and repeat ICER analyses. This package contains multiple dependencies to support syntax compatibility and plotting with ggplot2. 

## Installation

You can install the development version of easyicer with:

```         
install.packages("devtools")
devtools::install_github("stat545ubc-2024/easyicer", ref = "1.2.0")
```

## Usage

Here is a basic example of using this package with the `palmerpenguins::penguins` dataset:

```{r example}
library(easyicer)
library(palmerpenguins)

# Create a filter condition and store as an object
sex_male_year_2007 <- palmerpenguins::penguins$sex == "male" & palmerpenguins::penguins$year == 2007

summed_output(data = palmerpenguins::penguins,
              group_vars = c("island", "species"),
              sum_var = "bill_depth_mm",
              filter_vars = sex_male_year_2007,
              na.rm = TRUE)



```
