---
title: "Transforming raw data into ICER plots using the easyicer package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transforming raw data into ICER plots using the easyicer package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(easyicer)
library(here)
```

### Introduction

This vignette is a tutorial on how to seamlessly create incremental cost-effectiveness ratio (ICER) analyses, including plots, by making use of all the functions in this package. This package was designed in a pipeline structure so that the end-user could cohesively deploy the package functions in a step-wise manner to arrive at a final cost-effectiveness analysis (CEA) with ICER plots. For instructive purposes, this vignette will be organized into 3 use-cases that can be integrated as a single workflow or deployed standalone. The target domain for this package is health technology assessment and economics, where CEAs are a common tool for evaluating the risk-benefit of new health interventions for implementation.

### Data

The data used in this tutorial is a modified dataset obtained from the authors' research project. The full dataset is not publicly available. The modified dataset used herein can be located in the [GitHub](https://github.com/stat545ubc-2024/easyicer) repository for this package. For the purpose of this tutorial, we will assign the dataset to an object.

```{r}
# Load the "here" package to access the .csv file in the data folder if you are in the source package repository.
easyicer_data <- read.csv(here::here("data/easyicer_example_data.csv"))

dim(easyicer_data)

head(easyicer_data, n = 10)
    
```

The dataset contains 8 variables with 828 rows of observations. This data was obtained from a microsimulation model projecting lifetime health and economic outcomes of average-risk colorectal cancer screening simulations in pre-defined Canadian birth cohorts. There are 4 predictor variables, or characteristics (cohort, scr_age, ppt_rate, scr_mod, year), and 3 outcome variables (including both cost outcomes and a health outcome). See below for full variable definitions.

------------------------------------------------------------------------

### Variable Dictionary

1.  cohort: A simulated birth cohort in the Canadian population with modeled adenoma risk relative to the age group. Example: 1973 represents a cohort of individuals born between 1973 and 1977.

2.  scr_age: The modeled age of screening initiation for each scenario. Example: 50 means that simulated individuals began screening at age 50.

3.  ppt_rate: The modeled participation rate for colorectal cancer screening in the screening scenario. Example: 0.18 = 18% of the cohort underwent screening.

4.  scr_mod: The modality of the screening test. Either a fecal immunochemical test (FIT) or a colonoscopy (f10q).

5.  year: The corresponding year that outcomes were simulated for.

6.  Cost_of_screen_x_reen_positives: The total cost of screening, including follow-up screening of positive cases.

7.  Cost_of_clinic_x_ow_up_protocol: The cumulative cost incurred by year from clinical diagnoses and follow-up procedures.

8.  Health_adjusted_person_years: The estimated cumulative life years of all individuals based on individually ascribed health utility scores. Herein defined as quality-adjusted life years (QALY).

------------------------------------------------------------------------

### Use-Case 1: Calculating cumulative outputs for cost and QALY outcomes with `summed_output()`

Often times with health economic data, we might get a large volume of raw and unprocessed outputs. This is particularly the case for simulation models that generate and display highly granular information. Often times, the individual outputs are meaningless unless they are transformed in some way (e.g., calculating a cumulative value) This is where it is helpful to deploy the `summed_output()` function, especially when this process is iterative. The `summed_output()` function calculates the sum of a numeric column in the dataset to arrive at a cumulative value. With this dataset, this is valuable because we need to calculate the sum of each output variable (costs and QALY) since they were originally calculated per year. This function also allows the end-user to group by and filter specific characteristics. For example, in our data, we have multiple birth cohorts, screening ages and screening modalities that we want to group by before calculating cumulative values. Lastly, the function can handle any NA values that might otherwise disrupt the function calculation.

```{r}

# Set a filter parameter to call in filter_vars
ppt18 <- easyicer_data$ppt_rate == 0.18

# Calculate the cumulative values for the cost outcome
cumulative_cost <- summed_output(
  data = easyicer_data,
  group_vars = c("scr_mod", "scr_age"),
  sum_var = "Cost_of_screen_x_reen_positives",
  filter_vars = ppt18
)

# Calculate the cumulative values for the health outcome
cumulative_qaly <- summed_output(
  data = easyicer_data,
  group_vars = c("scr_mod", "scr_age"),
  sum_var = "Health_adjusted_person_years",
  filter_vars = ppt18
)

print(cumulative_cost)
print(cumulative_qaly)

```

Now we have arrived at our necessary outputs to calculate ICERs. Did you notice that this function is a general sum function with the ability to sum by groups or filtered datasets? This function is not strictly intended for calculating ICERs and is versatile with other purposes, but it makes iterative sum calculations very convenient - especially when calculating ICERs. Note that we did not specifically call `na.rm = TRUE`, this is because it is set to `TRUE` as default so if it is not called in the function it will automatically remove NA values.

------------------------------------------------------------------------

### Use-Case 2: Calculating ICERs from cumulative cost and QALY inputs with `icercalc()`

Before we can calculate ICERs, we need to transform the data structure so that the values are recognized in the correct order by the `icercalc()` function. This is because with ICERs, there are reference values that correspond with a reference or comparator scenario/intervention. In the context of our dataset, this refers to `scr_age = 50` and `scr_mod = FIT` (FIT 50), which is the current colorectal cancer screening strategy in Canada. In order to properly calculate the ICERs, the function expects a matrix structure as illustrated:

```         
        ref     new
 cost   n00     n01
 qaly   n10     n11
```

Once you have created a matrix, you might realize that the order of values in the matrix is not correct. Instead of manually re-coding the matrix, you can specify `rev =` as `"neither"`, `"rows"`, `"columns"` or `"both"` to reverse the order of the columns, rows, or columns and rows. By default `rev = "neither"` so you do not need to specify the `rev` argument if the matrix is correct. For instructive purposes, the `rev` argument will be showcased below.

```{r}

# Create matrix of cost and QALY values for each scenario
icer_matrix_fit45 <- matrix(c(
  cumulative_cost$cumulative[2],
  cumulative_cost$cumulative[1],
  cumulative_qaly$cumulative[2],
  cumulative_qaly$cumulative[1]),
  nrow = 2,
  ncol = 2,
  byrow = TRUE)

# Calculate ICERs using matrix 
icer_fit45 <- icercalc(data = icer_matrix_fit45)

# Re-iteration for COL 45 scenario
icer_matrix_col45 <- matrix(c(
  cumulative_cost$cumulative[2],
  cumulative_cost$cumulative[3],
  cumulative_qaly$cumulative[2],
  cumulative_qaly$cumulative[3]),
  nrow = 2,
  ncol = 2,
  byrow = TRUE)

icer_col45 <- icercalc(data = icer_matrix_col45)

# # Re-iteration for COL 50 scenario
icer_matrix_col50 <- matrix(c(
  cumulative_cost$cumulative[4], # The columns of the matrix are purposefully in inverse order to demonstrate rev argument.
  cumulative_cost$cumulative[2],
  cumulative_qaly$cumulative[4],
  cumulative_qaly$cumulative[2]),
  nrow = 2,
  ncol = 2,
  byrow = TRUE)

icer_col50 <- icercalc(data = icer_matrix_col50, rev = "columns")
```

The output of `icercalc()` displays the final matrix used in the ICER calculation, the incremental cost output (difference in cost between the new and reference scenarios), the incremental QALY (difference in QALY between new and reference scenarios) and the ICER. The component cost and QALY outcomes are displayed because if the end user works with the entire easyicer pipeline, both values will be required for the `icerplot()` function.

------------------------------------------------------------------------

### Use-Case 3: Creating ICER plots from calculated ICERs with `icerplot()`

The last stage of a basic CEA is to plot the ICERs along a coordinate system. This is helpful for visually comparing how cost-effective multiple interventions are relative to the reference intervention. The `icerplot()` function creates a scatterplot based on a dataframe with at least two numeric columns (representing cost and QALY outcomes). An additional and optional `names` argument is implemented which creates legend labels if the ICER values have corresponding names.

```{r}
# Create a dataframe (if needed) with cost and QALY derivatives of ICER values.
icer_coordinates <- data.frame(
intervention = c("FIT 45", "COL 45", "COL 50"),
QALY = c(17807.78, 86190.82, 77119.96),
Cost = c(176878878, 902374863, 794957011))

icerplot(data = icer_coordinates, x = "QALY", y = "Cost", names = "intervention")


```

Notice that you do not need the ICER outputs from `icercalc()` to use this function. In fact, the ICER values calculated with `icercalc()` will not work since both the component costs and QALY outcomes are required in the dataframe, which is why `icercalc()` provides both incremental costs and QALYs as well as the ICER. This highlights that these functions can be used in an integrated manner or independently based on the end user's needs and data structures.
